### 2017-10-13 Regeneron Paper.tex
Plenty of changes

\documentclass{article}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{blindtext}
\usepackage{enumitem}

\algnewcommand{\LineComment}[1]{
	\Statex \(\triangleright\) #1
}

\newcommand{\head}{
	\ensuremath{Head_L}
}
\newcommand{\tail}{
	\ensuremath{Tail_L}
}
\newcommand{\size}{
	\ensuremath{Size_L}
}
\newcommand{\capacity}{
	\ensuremath{Cap_L}
}

\newcommand{\headsize}{
	\ensuremath{Hsize_L}
}
\newcommand{\headcapacity}{
	\ensuremath{Hcap_L}
}
\newcommand{\isempty}{
	\ensuremath{Empty?_L}
}
\newcommand{\isfull}{
	\ensuremath{Full?_L}
}

\newcommand{\addfn}{
	\ensuremath{Add_L}
}
\newcommand{\resizefn}{
	\ensuremath{Resize_L}
}

\newcommand{\timeof}[1]{
	\ensuremath{T\big[#1\big]}
}
\newcommand{\spaceof}[1]{
\ensuremath{S\big[#1\big]}
}
\newcommand{\capacities}{
	\ensuremath{Caps_L}
}

\algrenewcommand{\textproc}{
}

\setlength{\parskip}{1em}
\setlist{nolistsep}

\begin{document}
	\begin{abstract}
		This paper introduces the \textit{fragmented list} data structure, which implements the list abstract data type. A fragmented list consists of multiple arrays, or \textit{buffers}, that store its elements. Fragmented lists do not make copies or throw away old buffers while resizing. As a result, they perform significantly better than dynamic arrays when a large number of elements are added to them. However, this is provided the final size of the list (or a close upper bound on it) is not known beforehand. If it is, a buffer of that size can be pre-allocated and resizing can be avoided, making fragmented lists redundant.
	\end{abstract}

	\section{Introduction}
	\label{Introduction}

	% TODO: This isn't being indented
	Often, node-based structures receive the most attention in a course on data structures. Although it is simpler to implement everything with arrays\footnote{}, node-based structures are favored because they have better time complexity for operations such as insertion and deletion.

	However, arrays have their own performance advantages over such structures, including:

	\begin{itemize}
		\item Better locality when iterating: elements of an array live next to each other in memory.
		\item Less memory overhead: memory does not need to be allocated for nodes.
		\item Fast random access: since arrays are just pointers, indexing them takes a few instructions and runs in $O(1)$ time.
	\end{itemize}

	When an abstract data type (which I will refer to as \textit{data type}) can be implemented easily with arrays, they are often used. One such data type is the \textit{list}\footnote{}, which is usually implemented with a dynamic array.

	Dynamic arrays perform reasonably well with respect to their most common operations. Adding an item runs in amortized $O(1)$ time, indexing them runs in $O(1)$ time, and they can be iterated in $O(n)$ time.

	However, when a dynamic array is \textit{full} (that is, its size equals its capacity) and an item is added, it must resize its buffer to fit the new item. Its behavior here is less than optimal: it allocates a new buffer larger than the current one, copies over the items, and discards the current buffer.

	Instead of throwing away buffers once they're filled up, \textit{fragmented lists} store references to them in a two-dimensional array. During a resize, a new buffer is still allocated; however, for a sufficiently large number of items, it is always half the size of the buffer that a dynamic array would allocate.

	In addition, no items are copied to the front of the new buffer. This is because no buffer represents the entire content of the list; each buffer holds just a \textit{fragment} of the items. To get back the items that were added to the fragmented list, one needs to traverse each of the buffers and copy their share of the items.

	% TODO: ToC or not?

	\section{Definitions}
	\label{Definitions}

	\subsection{Terminology}

	\begin{description}
		\item[buffer] An array that holds a portion of the items in a fragmented list.
		\item[capacity] The maximum amount of items a list can hold without resizing.
		\item[full, filled] A list is full iff its size equals its capacity, meaning it must resize if another item is added.
		\item[growth factor] The factor by which the capacity grows when the list is resized, if the current capacity is nonzero.
		\item[list] The list data type. This term does not refer to a particular implementation of the data type, only the "interface."
		\item[resize] To increase a list's capacity once it is full.
		\item[threshold sizes] The set of sizes for which the list is full.
	\end{description}

	% TODO: Prove that the set of threshold sizes equals the set of capacities. Superset/subset argument.

	\subsection{Runtime-Provided Functions}

	I assume the following functions are provided by the runtime environment, so that I may use them without defining them beforehand.\par

	\begin{algorithm}\begin{algorithmic}
			\LineComment{Adds \textit{item} to dynamic array \textit{d}}
			\State $Add\ dyn_{d}(item)$
			\\
			\LineComment{Allocates and returns an array of length \textit{len}}
			\State $Array(len)$
			\\
			\LineComment{Allocates and returns an empty dynamic array}
			\State $Dynamic()$
			\\
			\LineComment{Returns the length of array \textit{a}}
			\State $Len_a$
		\end{algorithmic}\end{algorithm}

	\subsection{Mathematical Notation}

	\begin{description}
		\item[$T_m(n)$] Time it takes to allocate memory for an array of size $n$
	\end{description}

	\section{Fragmented List Structure}
	\label{Fragmented List Structure}

	Each fragmented list, denoted by $L$, is given four \textit{fields}: \head, \tail, \size, and \capacity. These fields may be read from and assigned to.

	\begin{itemize}
		\item \head is the \textit{head} of the fragmented list. It returns the buffer we are currently adding items to.
		\item \tail is the \textit{tail} of the list. It returns a list of buffers that have already been filled.
		\item \size is the list's \textit{size}. It returns the number of items in $L$.
		\item \capacity is the list's \textit{capacity}. It returns the maximum number of items $L$ can hold without resizing.
	\end{itemize}

	Below, I also define some auxiliary functions on $L$ (which may not be assigned to). $:=$ denotes a definition, as opposed to $=$ which checks equivalence. Functions that return boolean values are suffixed with ?.

	\begin{algorithm}\begin{algorithmic}
			\LineComment{Returns whether $L$ is empty}
			\State $\isempty := \size = 0$
			\\
			\LineComment{Returns whether $L$ is full}
			\State $\isfull := \size = \capacity$
			\\
			\LineComment{Returns the capacity of \head}
			\State $\headcapacity := Len_{\head}$
			\\
			\LineComment{Returns the size of \head}
			\State $\headsize := \size - \left(\capacity - \headcapacity\right)$
		\end{algorithmic}\end{algorithm}

	Before we call any functions with $L$, we need to make sure that it is properly initialized. (In an object-oriented language, this code would go in the constructor.)

	\begin{algorithm}\begin{algorithmic}
			\State $\head \gets Array(0)$
			\State $\tail \gets Dynamic()$
		\end{algorithmic}\end{algorithm}

	\section{Fragmented List Operations}
	\label{Fragmented List Operations}

	\subsection{Adding an Element}

	I begin with the implementation for \textit{Add} since it is the most common list operation\footnote{}. Its algorithm is very similar to that of a dynamic array: it resizes the list if it's full, then stores the item and increments the size. The head's size is also incremented since an item is added to it.

	\begin{algorithm}\begin{algorithmic}[1]
			\Procedure{$Add_L$}{$item$}
				\If{$\isfull$}
					\State $Resize_L()$
				\EndIf

				\State $\head[\headsize] \gets item$
				\State $\size \gets \size + 1$
			\EndProcedure
		\end{algorithmic}\end{algorithm}

	A fragmented list is resized quite a bit differently from a dynamic array, however. % TODO

	\begin{algorithm}\begin{algorithmic}[1]
			\Procedure{$Resize_L$}{}
				\Require
					\Statex $\isfull$
				\If{$\isempty$}
					\State $\head \gets Array(4)$
					\State $\capacity \gets 4$
					\State \Return
				\EndIf

				\State $Add\ dyn_{\tail}(\head)$
				\State $new\ cap \gets G(\headcapacity)$
				\State $\head \gets Array(new\ cap)$
				\State $\headsize \gets 0$
				\State $\capacity \gets \capacity + new\ cap$
			\EndProcedure
		\end{algorithmic}\end{algorithm}

	\begin{algorithm}\begin{algorithmic}[1]
			\LineComment{\textit{G} is the \textit{growth function}.}
			\LineComment{It decides the capacity of the next buffer based on the current buffer's capacity.}
			\Function{\textit{G}}{$cap$}
			\Require
			\Statex $cap \geq 4 \land \log_2{cap} \in \mathbb{N}$
			\State \Return $\begin{cases}
			4 & cap = 4\\
			2 * cap & cap \neq 4
			\end{cases}$
			\EndFunction
		\end{algorithmic}\end{algorithm}

	\subsubsection{Time Analysis}

	In each "Time Analysis" section, I derive not just the time complexity of the operation, but a complete formula that approximates the actual (amortized) amount of time it takes. In order to do this in a clean fashion, I use the following notation:

	\begin{description}
		\item[$\timeof{F}$] The amortized amount of time it takes for function $F$ to run.
		\item[$T_A$] The amortized amount of time it takes for the code at line to run.
		\item[$T_{A,B}$] The amortized amount of time it takes for code between lines $A$ and $B$ to run, inclusive.
		\item[$T(k)$] The amount of time it takes for $T$ to run, where $T$ is any time expression, when $\size = k$.
		% TODO: T_m should go here?
	\end{description}

	Since the first three definitions represent \textit{amortized} time, which is the average amount of time it takes for an operation to run, I can expand them to

	\begin{align*}
	\timeof{F} &= \frac {\sum_{i = 0}^{n - 1} F(i)} n\\
	T_A &= \frac {\sum_{i = 0}^{n - 1} T_A(i)} n\\
	T_{A,B} &= \frac {\sum_{i = 0}^{n - 1} T_{A,B}(i)} n
	\end{align*}

	% TODO: Where n is an arbitrarily large integer? Since that's true, we may ignore 0 in the below sigmas.

	These definitions will be used frequently when a piece of code takes longer to run for some sizes than others.

	Now, I begin analyzing the amortized amount of time for \addfn. The expression we want to find is \timeof{\addfn}. In order to do this, I will need a few lemmas:

	Lemma. Let \capacities be the set of all values assigned to \capacity as \size increases indefinitely. Let $\capacities(n)$ be the subset of \capacities less than $n$. I show that $capacities(n)$ is the set

	\begin{align*}
	\capacities(n) = \{0\} \cup \{2^i | 1 < i \leq lgc(n) \}
	\end{align*}

	where $lgc(n) = \lfloor \log_2{(n - 1)} \rfloor$, the base 2 \textit{logarithm of the greatest capacity} less than $n$.

	% TODO: Proof

	We define $\overline{\capacities(n)}$ as $\{0, 1, ... n - 1\} \char`\\ \capacities(n)$.

	\begin{align*}
	\timeof{\addfn} &= T_{2,6} = T_2 + T_3 + T_{4,6}\\
	&= T_3 + C\\
	&= \frac {\sum_{i = 0}^n T_3(i)} n + C\\
	&= \frac {\sum_{i \in \capacities(n)} T_3(i) + \sum_{i \in \overline{\capacities(n)}} T_3(i)} n + C
	\end{align*}

	Since $|\capacities(n)| = lgc(n) - 1$ and $|\overline{\capacities(n)}| = n - |\capacities(n)| = n + 1 - lgc(n)$, this resolves to

	\begin{align*}
	\frac {(lgc(n) - 1)\timeof{\resizefn}} n + C
	\end{align*}

	Now we have to determine the complexity of \resizefn.

	\begin{align*}
	\timeof{\resizefn} &= T_{2, 10} = T{6,10}
	\end{align*}

	When $L$ is not full:

	$$
	T_{2,3} + T_{4,6} = \timeof{full?} + T_{4,6} = C
	$$

	When $L$ is full:

	$$
	T_{2,3} + T_{4,6} = \timeof{full?} + T_[resize] + T_{4,6} = T[resize] + C
	$$

	Now we must analyze $resize$.

	When $L$ is empty:

	$$
	T[resize] = T_{2,5} = T[empty?] + T_alloc(4) + T_5 = T_alloc(4) + C
	$$

	When $L$ is non-empty:

	$$
	T[resize] = T[empty?] + 
	$$

	\subsubsection{Space Analysis}

	\subsection{Indexing}

	\begin{algorithm}\begin{algorithmic}[1]
			\Function{\textit{item at index}}{$L, index$}
	 			\Require
	 				\Statex $0 \leq index < \size$
	 			\State $i \gets index$
	 			\ForAll{$buf \in \tail$}
	 				\If{$i < Len_{buf}$}
	 					\State \Return $buf[i]$
	 				\EndIf
	 				\State $i \gets i - Len_{buf}$
	 			\EndFor

	 			\Return $\head[i]$
 			\EndFunction
		\end{algorithmic}\end{algorithm}

	A function to set an item at a particular index can be implemented in a similar fashion.

	\subsection{Inserting an Element}

	\subsection{Deleting the Last Element}

	\subsection{Deleting an Element}

	\subsection{Copying to an Array}

	\subsection{Iterating}

	\subsection{Searching}

	\subsection{In-Place Sorting}

	\section{Implementations}
	\label{Implementations}

	\section{Benchmarks}
	\label{Benchmarks}

	\section{Closing Remarks}
	\label{Closing Remarks}

	% TODO: May need to ask something about how the pseudocode is only valid for GC'd languages.

\end{document}