### 2017-10-15 Draft2.tex
Revert some things from previous commit

\documentclass{article}
\usepackage{amsmath}

\newcommand{\descriptn}{\textbf{Description}}
\newcommand{\dynarrayimpl}{\textbf{Dynamic array implementation}}
\newcommand{\funarrayimpl}{\textbf{Funnel array implementation}}
\newcommand{\tcomplex}{\textbf{Time complexity}}
\newcommand{\scomplex}{\textbf{Space complexity}}
\newcommand{\tcomplexcmp}{\textbf{Time complexity comparison}}
\newcommand{\scomplexcmp}{\textbf{Space complexity comparison}}

\newcommand{\timefn}{T}
\newcommand{\spacefn}{S}
\newcommand{\nwritesfn}{W}

\newcommand{\timenewfn}{\timefn'}
\newcommand{\spacenewfn}{\spacefn'}
\newcommand{\nwritesnewfn}{\nwritesfn'}

\newcommand{\timeratio}{r_\timefn}
\newcommand{\spaceratio}{r_\spacefn}

\newcommand{\bigo}{O}
\newcommand{\biggo}{P}
\newcommand{\varnitems}{n}
\newcommand{\indexertime}{I}

\newcommand{\listname}{L}

\setlength{\parskip}{1em}

\begin{document}
	\begin{abstract}
	\end{abstract}

	\section{Introduction}

	\section{Definitions}

	\begin{description}
		\item[field] A variable associated with an object that may be read from or written to.\\
			It is assumed that fields can be read from and written to in constant time.
		\item[get indexer] A method that gets an item at a specified index in a list.
		\item[indexer] Refers to a get indexer or set indexer.
		\item[property] A function of an object that returns a value, runs in constant time, and does not mutate any state (i.e. by assigning new values to fields).
		\item[set indexer] A method that sets an item at a specified index in a list.
		\item[trivial member] A field or property of an object.
	\end{description}

	\section{Runtime-Defined Functions}

	\section{Trivial Members of Dynamic and Funnel Arrays}

	In this section, I define \textbf{fields} and \textbf{properties} on dynamic and funnel arrays. These will be used later to implement operations on them in pseudocode.

	\section{Common Funnel Array Operations}

	In this section, I implement common operations for dynamic and funnel arrays in pseudocode. Then, I analyze and compare the time complexities of the implementations. If the operation allocates memory, I also compare their space complexities.

	The following mathematical definitions will be used while analyzing time and space complexity:

	\begin{description}
		\item[$\biggo(f(n))$] This is an alternative to big-O notation for time complexity; I will call it \textbf{big-P notation}.
			\subitem The main purpose of big-P notation is to preserve the constant coefficient of the fastest-growing term in $f(n)$, while still maintaining many properties of big-O. For example, $\bigo(2n) = \bigo(n)$, but $\biggo(2n) \neq \biggo(n)$. This makes it possible to compare time complexities whose ratio converges to a constant as $n$ becomes larger.
			\subitem More formally, $\biggo(f(n)) = \biggo(g(n))$ iff $$\lim_{n \to \infty} {\frac{f(n)}{g(n)}} = 1$$.
	\end{description}

	\subsection{Adding}

	\descriptn

	Adding is the most common operation done on a dynamic array\footnote{}. Funnel arrays improve the performance of adding in two ways: by allocating less memory, and reducing the amount of copying.

	\dynarrayimpl

	\tcomplex

	\begin{align*}
	\nwritesfn(\varnitems) = \biggo(2\varnitems)
	\end{align*}

	\scomplex

	\begin{align*}
	\spacefn(\varnitems) = \biggo(2^{\lceil \log_2 \varnitems \rceil + 1}) = \bigo(\varnitems)
	\end{align*}

	\funarrayimpl

	\tcomplex

	\begin{align*}
	\nwritesnewfn(\varnitems) = \biggo(\varnitems)
	\end{align*}

	\scomplex

	\begin{align*}
	\spacenewfn(\varnitems) = \biggo(2^{\lceil \log_2 \varnitems \rceil}) = \bigo(\varnitems)
	\end{align*}

	\tcomplexcmp

	\begin{align*}
	\end{align*}

	\scomplexcmp

	\begin{align*}
	\spaceratio = \frac {\biggo(2^{\lceil \log_2 \varnitems \rceil})} {\biggo(2^{\lceil \log_2 \varnitems \rceil + 1})} = \frac{1}{2}
	\end{align*}

	\subsection{Indexing}

	\descriptn

	Indexing is another very common operation on a list. An \textbf{indexer} gets or sets an item at a specified index in a list.

	\dynarrayimpl

	\tcomplex

	$\timefn(\varnitems) = \bigo(1)$

	% <funnel array implementation>

	\tcomplex

	$\timenewfn(\varnitems) = \bigo(1)$

	\tcomplexcmp

	\subsection{Iterating}

	\descriptn

	\textbf{Iteration} of a list is the process of performing some action on each of its elements.

	\dynarrayimpl

	\tcomplex

	Ignoring the arbitrary code run after getting each element, the time complexity of this method is $\bigo(\varnitems)$.

	\funarrayimpl

	\tcomplex

	$\timenewfn(\varnitems) = \bigo(\varnitems)$

	\tcomplexcmp

	\subsection{Copying to an array}

	\section{Other Funnel Array Operations}

	\subsection{Inserting}

	\subsection{Deleting}

	\subsection{Searching}

	\subsection{In-place sorting}

	\section{Implementations}

	\section{Benchmarks}

	\section{Closing Remarks}

\end{document}