### 2017-10-17 Draft2/Abstract.tex
Add the abstract

This paper introduces the \textbf{growth array}, an array-like data structure designed to be efficient at appending elements. They offer substantial time and memory savings over dynamic arrays when a large number of elements are appended. Growth arrays also support other list operations, such as random access, iteration, and insertion/deletion at an index. However, they perform no better, or slightly worse, than dynamic arrays for those methods. Also, if the size of the data is known beforehand, allocating a raw array with that size and appending to it is faster than appending to a growth array.