### 2017-10-18 Draft2/CommonOperations.Blurb.tex

In this section, I implement common operations for dynamic and growth arrays, and analyze their time complexity. I also analyze space complexity of operations that allocate memory.