### 2017-10-24 Draft2/Abstract.tex
Revise Intro, Abstract

This paper introduces \textbf{growth arrays}, array-like data structures that are designed for appending elements. When the number of items is not known beforehand but is expected to be large, they are more efficient than dynamic arrays by a constant factor. Growth arrays support all operations dynamic arrays do, such as random access, iteration, and insertion or deletion at an index. However, they perform no better, or slightly worse, than dynamic arrays for operations other than appending.