### 2017-10-25 Draft2/CommonOperations.Blurb.tex
Get the theorems straight?

In this section, I implement the most common operations for dynamic and growth arrays, then analyze their time complexity. If the operation allocates memory, I also analyze its space complexity.