### 2017-11-08 Draft2/Indexing.Dynamic.tex

\HdrDynArrayImpl

The following algorithm implements $\FuncGetItem$ for dynamic arrays. A $\FuncSetItem$ function can be implemented in a similar fashion.

{\HdrNote} I assume all arguments passed to my algorithms are valid. Thus, I do not include argument validation nor error handling in my pseudocode.

\begin{algorithm}
	\begin{algorithmic}[1]
		\Function{$\FuncGetItem$}{$\VarList, index$}
			\State \Return $\VarList.\FieldBuffer[index]$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

Clearly, this function runs in constant time. It also has the benefit of using very few instructions: on architectures like x86, indexing the buffer may take as little as one instruction.