### 2017-11-11 Draft2/CopyArray.Blurb.tex
Finally finish the Common Ops section

% Todo: Is 'n memory' vague?
After they are done appending to them, users often want to take dynamic arrays and convert them into raw arrays. There are multiple possible reasons for this: 1) Raw arrays hold on to exactly $n$ memory to store $n$ elements. However, dynamic and growth arrays typically use $> n$ memory so they need not grow every time $\FuncAppend$ is called. 2) A function in third-party code might only accept a raw array as an argument.

The case is even more compelling for growth arrays: 3) Since they are not entirely contiguous, they have worse data locality (even if the number of discontinuities is $O(\log n)$). 4) Raw arrays can be indexed much faster than growth arrays.