### 2017-11-11 Draft2/CopyArray.Dynamic.tex

\HdrDynArrayImpl

The algorithm for converting a dynamic array to a raw array is straightforward:

\begin{algorithm}
	\begin{algorithmic}
		\Function{$\FuncToArray$}{$\VarList$}
			\State $\LclRawArray \gets \FuncNewArray(\VarList.\FieldSize)$
			\State $\FuncArrayCopy(\VarList.\FieldBuffer,\ \LclRawArray,\ \VarList.\FieldSize)$
			\State \Return $\LclRawArray$
		\EndFunction
	\end{algorithmic}
\end{algorithm}