### 2017-11-11 Draft2/Iterating.Dynamic.tex

\HdrDynArrayImpl

The algorithm for iterating dynamic arrays is simple. It loops through all valid indices, and calls the $\FuncGetItem$ method (which is denoted $\VarList[i]$ instead of $\VarList.\FuncGetItem(i)$) for each index.

\begin{algorithm}
	\begin{algorithmic}
		\LineComment{Algorithm for ``$\TextFor \LclItem \TextIn \VarList$'', where $\VarList$ is a dynamic array}
		\For{$\LclIndex = 0,\ \LclIndex < \VarList.\FieldSize,\ \LclIndex \gets \LclIndex + 1$}
			\State $\LclItem \gets \VarList[\LclIndex]$
			\State $\ParamAction(\LclItem)$
		\EndFor
	\end{algorithmic}
\end{algorithm}