### 2017-11-13 Draft2/CommonOperations.Blurb.tex

In this section, I implement commonly used operations for dynamic and growth arrays, then analyze their time complexity. I also analyze the space complexity of operations that allocate memory.