### 2017-11-13 Draft2/Implementations.tex

% Todo: Use hyperref?
In order to demonstrate empirical evidence that the algorithms described above work, I have implemented them in the programming language C\#. I have written a full test suite for them and benchmarked them extensively. The implementation and the benchmark results can be found here: https://github.com/jamesqo/StsProject

I have previously implemented this algorithm for the open-source project .NET Core, which is run by Microsoft. I submitted two patches that applied this algorithm in order to reduce memory consumption for a commonly-called function. The patches can be found at the following links: https://github.com/dotnet/corefx/pull/11208 and https://github.com/dotnet/corefx/pull/13076. The first patch contains the original implementation of the algorithm. The second patch contains a cleaned-up version of the original implementation.