### 2017-11-13 Draft2/Abstract.tex
[WIP] Feedback

This paper introduces \textbf{growth arrays}, array-like data structures that are designed for appending elements. When the number of items is not known beforehand but is expected to be large, growth arrays are more efficient than dynamic arrays by a constant factor. They support all operations dynamic arrays do, such as random access, iteration, and insertion or deletion at an index. For operations other than appending, however, they perform the same as or slightly worse than dynamic arrays.