### 2017-11-14 Draft2/CopyArray.Blurb.tex
Work

Users often want to convert dynamic arrays to raw arrays. There are multiple reasons for this: 1) Raw arrays hold on to exactly $n$ memory to store $n$ elements. However, dynamic and growth arrays typically use more than $n$ memory, so they need not grow every time an item is appended. 2) Sometimes, functions in third-party code only accept raw arrays.

For growth arrays, the case is even more compelling: 3) Since they are not contiguous as a whole, they have worse data locality (even if the number of discontinuities is $O(\log n)$). 4) Raw arrays can be indexed much faster than growth arrays.

In this section, I will show that growth arrays can be converted to raw arrays not much slower than dynamic arrays.