### 2017-11-14 Draft2/CopyArray.Dynamic.tex

\HdrDynArrayImpl

It is straightforward to convert a dynamic array to a raw array:

\begin{algorithm}[H]
	\caption{Converting to raw array \TextDynamicArray}
	\begin{algorithmic}
		\Function{$\FuncToArray$}{$\VarList$}
			\State $\LclRawArray \gets \FuncNewArray(\VarList.\FieldSize)$
			\State $\FuncArrayCopy(\VarList.\FieldBuffer,\ \LclRawArray,\ \VarList.\FieldSize)$
			\State \Return $\LclRawArray$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

Since $n$ elements must be copied, the time complexity of this function is $O(n)$.