### 2017-11-14 Draft2/CopyArray.Growth.tex

\HdrGrowthArrayImpl

The growth array algorithm is slightly slower than the dynamic array one. It needs to loop over the tail and call $\FuncArrayCopy$ multiple times, since not all elements are contiguous. However, $n$ elements are still copied in total, so the time complexity is still $O(n)$.

\begin{algorithm}[H]
	\begin{algorithmic}
		\Function{$\FuncToArray$}{$\VarList$}
		\State $\LclRawArray \gets \FuncNewArray(\VarList.\FieldSize)$
		\For{$\LclBuffer \TextIn \VarList.\FieldTail$}
			\State $\FuncArrayCopy(\LclBuffer,\ \LclRawArray,\ \LclBuffer.\FieldLength)$
		\EndFor
		\State $\FuncArrayCopy(\VarList.\FieldHead,\ \LclRawArray,\ \VarList.\FieldHeadSize)$
		\State \Return $\LclRawArray$
		\EndFunction
	\end{algorithmic}
\end{algorithm}