### 2017-11-14 Draft2/Indexing.Dynamic.tex

\HdrDynArrayImpl

The following algorithm implements $\FuncGetItem$ for dynamic arrays. (A $\FuncSetItem$ function can be implemented in a similar fashion.)

{\HdrNote} My algorithms assume all arguments passed to them are valid. Thus, arguments are never checked.

{\HdrNote} Although the function is called $\FuncGetItem$, the syntax other algorithms use to call it is $\VarList[\ParamIndex]$.

\begin{algorithm}[H]
	\caption{Random access \TextDynamicArray}
	\begin{algorithmic}[1]
		\Function{$\FuncGetItem$}{$\VarList,\ \ParamIndex$}
			\State \Return $\VarList.\FieldBuffer[\ParamIndex]$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

This algorithm runs in $O(1)$ time. It also has the benefit of using very few instructions, which is a benefit that will not be shared with growth arrays.