### 2017-11-14 Draft2/Iterating.Dynamic.tex

\HdrDynArrayImpl

The algorithm for iterating dynamic arrays is simple and familiar.

\begin{algorithm}[H]
	\caption{``$\TextFor \LclItem \TextIn \VarList$'' algorithm \TextDynamicArray}
	\begin{algorithmic}
		\For{$\LclIndex = 0,\ \LclIndex < \VarList.\FieldSize,\ \LclIndex \gets \LclIndex + 1$}
			\State $\LclItem \gets \VarList[\LclIndex]$
			\State $\TextDoSomethingWith \LclItem$
		\EndFor
	\end{algorithmic}
\end{algorithm}