### 2017-11-14 Draft2/Iterating.Growth.tex

\HdrGrowthArrayImpl

The iteration algorithm for growth arrays is not quite as simple. As shown in the previous section, growth arrays' random access algorithm is expensive; thus, it is best avoided. The following algorithm implements iteration without using the random access algorithm.

\begin{algorithm}[H]
	\caption{``$\TextFor \LclItem \TextIn \VarList$'' algorithm \TextGrowthArray}
	\begin{algorithmic}
		\For{$\LclBuffer \TextIn \VarList.\FieldTail$}
			\For{$\LclItem \TextIn \LclBuffer$}
				\State $\TextDoSomethingWith \LclItem$
			\EndFor
		\EndFor
		\For{$\LclIndex = 0,\ \LclIndex < \VarList.\FieldHeadSize,\ \LclIndex \gets \LclIndex + 1$}
			\State $\LclItem \gets \VarList.\FieldHead[\LclIndex]$
			\State $\TextDoSomethingWith \LclItem$
		\EndFor
	\end{algorithmic}
\end{algorithm}