### 2017-11-15 Draft2/ExternalDefinitions.tex

I assume that the following functions are defined by the runtime. Thus, I will use them in my algorithms without providing definitions for them.

\begin{algorithm}[H]
	\caption{Runtime-defined functions}
	\begin{algorithmic}
		\LineComment{Copies $len$ items from $source$ to $dest$}
		\State $\FuncArrayCopy(source,\ dest,\ len)$
		\State
		\LineComment{Returns a new array with length $len$}
		\State $\FuncNewArray(len)$
		\State
		\LineComment{Returns a new, empty dynamic array}
		\State $\FuncNewDynamicArray()$
	\end{algorithmic}
\end{algorithm}

\subsection{Constants}

I will use the following constants in my algorithms. These constants govern how the algorithms will behave. Specific values for them must be supplied by the person using the algorithms.

\begin{description}
	\item[initial capacity] Denoted by $\VarInitCapacity$; the capacity of an empty dynamic array. $\VarInitCapacity$ must be a natural number.
	\item[growth factor] Denoted by $\VarGrowthFactor$; the constant ratio of the new capacity to the old one when $\VarList$ grows. $\VarGrowthFactor$ must be greater than $1$.
\end{description}