### 2017-11-15 Draft2/Indexing.Dynamic.tex
Prune indexing section

\HdrDynArrayImpl

The algorithm for indexing a dynamic array is trivial, and will not be shown here. It simply reads or writes to the $i$th element of $\VarList.\FieldBuffer$ (where $i$ is zero-based). It is well-known that this algorithm runs in $O(1)$ time. It also has the benefit of using very few instructions, which will not be the case for growth arrays' algorithm.