### 2017-11-15 Draft2/Iterating.Dynamic.tex
Work

\HdrDynArrayImpl

The algorithm for iterating dynamic arrays is well-known: let $i$ range from $0$ to $n - 1$ and, for each $i$, do something with $\VarList[i]$. Ignoring what is done with each item, the time complexity of this operation is $O(n)$.