### 2017-11-15 Draft2/Iterating.Growth.tex

\HdrGrowthArrayImpl

The iteration algorithm for growth arrays is not quite as simple. As shown in Section \ref{subsec:Indexing}, growth arrays' random access algorithm can be expensive; thus, it is best avoided. The following algorithm implements iteration without using the random access algorithm.

\begin{algorithm}[H]
	\caption{``$\TextFor \LclItem \TextIn \VarList$'' algorithm \TextGrowthArray}
	\begin{algorithmic}
		\For{$\LclBuffer \TextIn \VarList.\FieldTail$}
			\For{$\LclItem \TextIn \LclBuffer$}
				\State $\TextDoSomethingWith \LclItem$
			\EndFor
		\EndFor
		\For{$\LclIndex = 0,\ \LclIndex < \VarList.\FieldHeadSize,\ \LclIndex \gets \LclIndex + 1$}
			\State $\TextDoSomethingWith \VarList.\FieldHead[\LclIndex]$
		\EndFor
	\end{algorithmic}
\end{algorithm}

Since this algorithm loops over $n$ items, its time complexity is $O(n)$. There is nothing to suggest that this algorithm should be substantially slower than dynamic arrays' algorithm. Even though the inner loop is occasionally `interrupted' by having to switch buffers, the number of such interruptions is $O(\log n)$, so their cost is minuscule. Thus, the only substantial work done by this algorithm is iterating the buffers, which is roughly the same as the work done by dynamic arrays' algorithm.