### 2017-11-15 Draft2/Bibliography.tex
Bogus citation

\begin{thebibliography}{9}

\bibitem{ceillog2}
	Ko, James L. "Algorithm for implementing ceiling of log 2." \textit{GitHub}, 12 Nov. 2017,\\ \texttt{https://github.com/jamesqo/StsProject/blob/master/src/StsProject/Internal/MathHelpers.cs}
\end{thebibliography}