### 2017-12-11 Draft2/Bibliography.tex

\begin{flushleft}
\begin{thebibliography}{9}

\bibitem{tildedef}
	de Bruijn, Nicolaas G. \textit{Asymptotic Methods in Analysis}. Dover Publications, 1981.

\bibitem{littleodef}
	Cormen et al. \textit{Introduction to Algorithms, Second Edition}. The MIT Press, 2001.

\bibitem{amortized}
	Fiebrink, Rebecca. \textit{Amortized Analysis Explained}. Princeton University, 2007. Available at \texttt{https://www.cs.princeton.edu/~fiebrink/423/AmortizedAnalysisExplained\_Fiebrink.pdf}.

\bibitem{fptables}
	Fog, Agner. \textit{Lists of instruction latencies, throughputs and micro-operation breakdowns for Intel, AMD and VIA CPUs}. Technical University of Denmark, 2017. Available at \texttt{http://www.agner.org/optimize/instruction\_tables.pdf}.

\bibitem{ceillog2}
	Leiserson et al. \textit{Using de Bruijn sequences to Index a 1 in a Computer Word}. MIT Laboratory for Computer Science, 1998. Available at \texttt{http://supertech.csail.mit.edu/papers/debruijn.pdf}.

\end{thebibliography}
\end{flushleft}