### 2017-12-11 Draft2/Abstract.tex
Make a bunch more changes for NYCSEF

This paper introduces \textbf{growth arrays}, array-like data structures that are designed for appending elements. When the number of items is not known beforehand but is expected to be large, growth arrays perform better than dynamic arrays by a constant factor. They support all operations dynamic arrays do, such as random access, iteration, and insertion or deletion at an index. For operations other than appending, however, they perform the same as or slightly worse than dynamic arrays.

Growth arrays have the potential to make a major impact on the field of computer science. If dynamic arrays are replaced with growth arrays across various programming languages and frameworks, it is likely that many programs will see significant performance boosts. Also, other array-based data structures such as circular queues and stacks stand to benefit from the efficiency improvements of growth arrays.