### 2017-12-11 Draft2/ClosingRemarks.tex

In this paper, I formally developed the concept of growth arrays. By implementing several common dynamic array operations for them, I showed that they are a viable replacement for dynamic arrays. I proved that growth arrays outperform dynamic ones both time- and memory-wise when a large number of elements are appended. I exhibited logarithmic and constant time random access algorithms for growth arrays. Finally, I showed that they perform comparably to dynamic arrays for iteration and conversion to raw arrays.

Dynamic arrays are used frequently today because they are efficient at appending. Since growth arrays are even more efficient at it, they should be used in place of dynamic arrays when appending is performance-critical. Programming languages and frameworks should strongly consider adding support for growth arrays to improve program performance.

I will conclude this paper with several open-ended questions, which may serve as inspiration for further research.

\begin{itemize}
	\item Can other array-based data structures, such as stacks or circular queues, also benefit from using growth arrays' algorithm for growth?
	\item Is there a constant time random access algorithm for growth arrays that does not use floating-point and works for all values of $\VarGrowthFactor$ and $\VarInitCapacity$?
	\item To what extent does growth arrays' poorer data locality result in performance losses? (Can tuning memory allocators offset these losses?)
	\item How can binary search and sorting algorithms be best implemented for growth arrays? (In particular, growth arrays with $\VarGrowthFactor = 2$ have an evenly-divided structure: if $2^k$ items are appended, $2^{k - 1}$ will be in the head and $2^{k - 1}$ in the tail. Does this lend itself to a better binary search algorithm?)
\end{itemize}